### 2019-06-04 Documents/literature_review/pages/review_content.tex
Changes; - Final draft has been made for the re-write of the first lit review section

\section{Introduction}
The increasing use of public key cryptography by instant messaging and secure email means key fingerprint 
verification is an ever more important task. One of the biggest risks to the security of the communication channel is a Man-in-the-middle (MiTM) attack . A successful MiTH attack can circumvent the encryption as it allows an attacker to read all the encrypted data. A countermeasure for this is the verification of each parties' fingerprint. \textbf{[...]}

@@

% \textbf{TODO: } - Improve the flow of the first section, need to phrase it like I'm telling a story
% \textbf{TODO: } - Create written introduction setting the scene. \\
% \textbf{TODO: } - Add participant size to M. Shirvanian (25) \\
% \textbf{TODO: } - Create a background section for Compare-and-X
% \\
% \textbf{TODO: } - Create comparison area where inconsistencies are compared

\section{Literature Review}
\subsection{Fingerprint representation comparison}

Fingerprints are used to quickly compare the similarity of two keys. The comparison occurs between the digest of a strong one-way hash function. Historically these have been represented as a hexadecimal string and users were tasked with manually comparing fingerprints between two substrates, for example, a monitor screen and a business card. We will refer to this process throughout the paper as the \textit{``authentication ceremony''}. \\
It has been shown that the average human can only hold 7-digits worth of data in our working memory\cite{miller1956magical}. This rules out the possibility of comparing full digests. For examples SHA-1 is 40 hex digits (160-bit) and, therefore, will make it difficult for effective comparison. Therefore , if the  process is to involve humans interaction, there is a need for schemes that work effectively with our abilities and senses.

Research in this area has taken various encoding schemes and compared their fallibility to impersonated attacks. Impersonation of a key would involve brute force of the encoding schemes through the mass creation of keys. The main elements of comparison have been accuracy of attack detection and time to complete the comparison. In this context we will consider the metrics a measure of security and usability respectively.

\subsection*{Effective schemes}
Results 
from the literature consistently show the effectiveness of language based encodings (Words or Sentences). This was shown in \cite{dechand2016empirical}, \cite{tan2017can} and \cite{kainda2009usability} with the accuracy of these schemes all being above 94.00\%. In all cases these were the best scemes in terms of accuracy. The exception to this is the work performed by \textbf{Hsiao, \textit{et al.}}\cite{hsiao2009study} in 2009 with Words achieving an abnormal accuracy of 63.00\% (Comparisons of methodology will be discussed in later sections).
\\\\
Aside from textual representations were graphical schemes. Examples of schemes assessed were: Random Art\cite{perrig1999hash}, Flag\cite{ellison2003public}, T-Flag\cite{lin2010spate}, Vash\footnote{https://github.com/thevash/vash}, OpenSSH visual host key and Unicorns\footnote{https://unicornify.pictures/} to name a few. Accuracy from these schemes were mixed with positive results obtained by \textbf{Hsiao, \textit{et al.}}\cite{hsiao2009study} where their best accuracy was achived by Random Art (94.00\%) and Flag Ext.\footnote{The authors' own improvement on Flag}(88.00\%). On the other hand, the worst accuracy was obtained by Flag(50\%). The only other paper assessing graphical representations was the work of \textbf{Tan \textit{et al.}}\cite{tan2017can} where they also achieved mixed results with accuracies ranging from 46\% to 90\%.
\\
In terms of usability of graphical schemes the literature concurred on their ease-of-use. In terms of speed of comparison all schemes consistently obtained some of the quickest comparisons times. This is also supported indirectly from the the comparison results of alternative literature \cite{kainda2009usability}\cite{dechand2016empirical} where with these in consideration graphical schemes still remain the quickest overall.
In terms of research into the performance of graphical schemes, the literature does not contain an extensive review with only two papers including graphical schemes in their comparison. This is, therefore, a candidate for further research.

\subsection*{Experimentation comparison}

\subsection{Fingerprint representation schemes}
Another area of research is investigations into the actual physical encodings of the hash digest. This section will briefly discuss the current research available on the creation and security of actual encoding schemes. The actual details of the operation of the schemes are outside the scope of this literature review, therefore, minimal attention will be allocated to these details.

@@

\subsection{Fingerprint representation attacks}
This area of research studies ways to physically execute attacks on fingerprint encoding schemes. This differs from previously examined work due to papers discussing the performance and fallibility of encoding schemes simulated the attack without consideration for how the attack would be performed. Research in this area is scant, with lots of research attention being directed towards the security of Man-in-the-Middle (MITM) attacks and not the encoding schemes themselves.

>>>>

% \section{Introduction}
% The increasing use of public key cryptography by instant messaging and secure email means key fingerprint 
% verification is an ever more important task. One of the most significant risks to the security of the communication channel is a Man-in-the-middle (MiTM) attack. A successful MiTM attack can circumvent the encryption as it allows an attacker to read all the encrypted data. A countermeasure for this is the verification of each parties' fingerprint. \textbf{[...]}

@@

\section{Literature Review}

Fingerprints are used to compare the similarity of two keys quickly. The comparison occurs between two digests of a secure one-way hash function. Historically these have been represented as a hexadecimal string whereon verification fingerprints are compared between two substrates, for example, a monitor screen and a business card.
This process is known as the \textit{``authentication ceremony.''} \\
Prior research has shown that the average human can only hold around 7-digits worth of data in their working memory\cite{miller1956magical}. This rules out the possibility of comparing complete digests. For example, SHA-1 is 40 hex digits (160-bit) and, therefore, difficult to effectively compare. Therefore, if human interaction is required; there is a need for schemes that work effectively with consideration to individual limitations.

Research in this area has taken various encoding schemes and compared their fallibility to impersonated attacks. The main elements of comparison have been the ``accuracy of attack detection" and ``time to compare". In this paper, these are considered the metrics of ``security'' and ``usability'', respectively.

\subsection*{Definition of terms}
Throughout the paper, various terms are used to define elements of the authentication ceremony relevant to fingerprints verification. Explicit descriptions are, therefore, stated in this section. These are used to remove ambiguity and underlying associations readers may have for the words chosen.

\textbf{\textit{Encoding schemes}} - This is the physical method of encoding used to represent the fingerprint. For example, ``Hexadecimal'' or ``Words'' are a unique way to represent a fingerprint and thus, are encoding schemes.

\textbf{\textit{Method of comparison}} - This is how the user assesses the encoding scheme. The most common example and the one used as default is ``Compare-and-Confirm'' (CaC). CaC, as the name suggests, is the \textit{comparison} of two fingerprints on different devices or mediums that are then \textit{confirmed}. The following sections explain alternative methods of comparisons.

\subsection{Authentication ceremony performance}

\subsection*{Encoding scheme performance}
Results 
from the literature consistently show the effectiveness of language-based encodings such as Words or Sentences with accuracies ranging from 94\% \cite{dechand2016empirical}\cite{tan2017can}\cite{kainda2009usability}. In all cases, these were the best schemes from the sets assessed. The exception to this is the work performed by \textbf{Hsiao, \textit{et al.}}\cite{hsiao2009study} in 2009 with Words achieving an abnormal accuracy of 63.00\%.
\\\\
Aside from textual representations were graphical schemes. Examples of schemes assessed were: Random Art\cite{perrig1999hash}, Flag\cite{ellison2003public}, T-Flag\cite{lin2010spate}, Vash\footnote{https://github.com/thevash/vash}, OpenSSH visual host key and Unicorns\footnote{https://unicornify.pictures/}, among others. These schemes had mixed accuracy with ranges as large as 50\% - 94\% in work by \textbf{Hsiao, \textit{et al.}}\cite{hsiao2009study}. The only other paper assessing graphical representations was the work of \textbf{Tan \textit{et al.}}\cite{tan2017can} where they also achieved mixed results with accuracies ranging from 46\% to 90\%.
\\
In terms of usability of graphical schemes, the literature concurred on their high usability. The comparison speed of these schemes were all among the quickest (See Table \ref{tab:times} for an overview of timings). Other work also indirectly concurred with graphical encodings having significantly quicker comparison times compared to non-graphical schemes \cite{dechand2016empirical}\cite{kainda2009usability}.
In terms of research into the performance of graphical schemes, the literature does not contain an extensive review with literature only containing two papers. There is also no overlap in the schemes assessed with each paper reviewing a unique set. This is, therefore, a promising candidate for further research.

One unique paper in this research area was the work by \textbf{M. Shirvanian \textit{et al.}}\cite{shirvanian2017pitfalls} produced in the context of secure messaging pairing. This paper was unique for several reasons. First was to consideration for ``remote-vs-proximity'' pairing where this is the first consideration of this aspect found in the literature. There is room for further research to compare encoding schemes in the context of ``remote-vs-proximity.'' Another unique aspect was the end-to-end encryption context of the study.

The findings from the paper showed a high false negative rate for all the schemes; this is a consideration also missing from the literature. Alongside this, results for usability were lower in a remote setting for all the schemes. The author, however, comments on the expected nature of this result.\\
Images were assessed as being the most secure method of authentication in the remote setting but voted as the method with the worst usability. This conclusion is highly inconsistent with all other work in this area. However, this could be due to the unique setting of remote verification, resulting in distinct results from user studies. Without further work in this area, it is difficult to validate these results conclusively.

Alongside this, it was shown in work by \textbf{Hsiao, \textit{et al.}}\cite{hsiao2009study} that age and gender do not affect the accuracy of the scheme. However, younger participants were considerably faster. Furthermore, findings also showed that language comprehension helped in discerning small differences between schemes encoded in Chinese, Japanese, Korean or English. Subsequently, knowledge of the language did not assist in differentiating more significant changes in the schemes as these had high accuracy regardless. These were interesting and unique considerations. Further work could aim to corroborate these conclusions.

Tables~\ref{tab:results} \& \ref{tab:times} contain the accuracy results of all papers assessed; this is to aid in visual comparison. Each paper used a different metric for measuring accuracy. Therefore, all results have been translated into ``overall accuracy.''

\subsection*{Methods of comparison performance}
Aside 
from ``Compare-and-Confirm'' (CaC), there is ``Compare-and-Select'' (CaS) and ``Compare-and-Enter'' (CaE). CaS is the method where one device displays the fingerprint, and the other user is provided with several options. The user then has to choose the correct value from the list of candidates. If there is no match, the user must deny the connection attempt. The creation of CaS was due to concerns that CaC would be ``too easy'' for users leading to complacency and errors\cite{uzun2007usability}. CaE is designed for scenarios where both devices might not have a display, i.e. pairing between a phone and a keyboard. One device displays the checksum. This checksum is entered into the other device. The first device then compares the entered string and checks for a match.

Research has been performed assessing the performance of these schemes and how they affect the ultimate security of the authentication ceremony. The literature agrees on CaC being the best overall scheme to compare fingerprints \cite{tan2017can}\cite{uzun2007usability} with CaS being highlighted for its poor security and usability. CaE has  had contrasting results. \cite{uzun2007usability} discarded it after one round due to ``poor usability.'' However, \cite{tan2017can} considered it the best method overall for usability and security. These conflicting results, therefore, show polarisation in the results of CaE. However, this could be due to the different overall use-cases of the studies. Validation of results from either study would, therefore, be an area of additional study.

% One aspect of note from the work by \textbf{Ersin Uzun \textit{et al.}}\cite{uzun2007usability} is their studies context  of  “Secure  device  pairing”  and  thus  some comparison schemes are tested that are not relevant in a fingerprint comparison context, for example “Choose-and-Enter” (A participant chooses a passphrase from an available list and enters it into the other device).  Thus, only relevant comparison schemes and their results were extracted and compared. \\

\subsection*{Experimentation methodology comparison}
To further look into the validity of previously discussed results, it is necessary to asses how the respective studies reached these conclusions. Areas for consideration are scheme entropy, attacker strength and participant attributes.

The starkest limitations of the literature are the range of participants and encoding entropy. The worst studies tested only 22-bits of entropy. This makes it difficult to compare results directly. One of the papers this most effects is the early work by \textbf{Hsiao, \textit{et al.}}\cite{hsiao2009study}  where their highest entropy is 28-bits. This level of entropy was inadequate even at the time of publication. There is an attempt by the authors' to address this issue in the later stages of the paper where they write \textit{``[...] increasing entropy is not a solution because it sacrifices usability and accuracy. With more entropy, representations will contain longer sequences of characters or more minute details which will lead to increased time and errors during comparisons.''}. This statement is backed up with no empirical evidence, and the authors' fail to consider how the low entropy would affect the overall security of the schemes.

Attacker strength is also another metric used to compare results concluded by each paper. Some papers failed to address their attacker strength consideration directly, but the overall strength has been inferred from the changes made to their schemes. For example, if they decided to change a single character in a 40-digit (160-bit) SHA-1 hex digest, they are indirectly stating that the attacker can control 39-digits (156-bit). To achieve this, the attacker would have to compute $2^{156}$ SHA-1 compressions to find a key match.
In the literature, this element ranged from $2^{28}$ to around $2^{242}$; however, this has some relation to the size of the encoding schemes used. This is a substantial range that makes it ultimately challenging to compare and confer results confidently. Further work could exclusively look into the affects attacker strength has on the success of attacks. Moreover, all the papers assessed failed to fully consider the feasibility of attacks in terms of computer and storage requirements. This again, shows gaps in the literature.

All of the studies considered demographical data when presenting their results. Their average ages were all around $\sim35$ years old with the majority of participants educated with at least a bachelor degree. This was alongside the equal split between male and female participants.

One consideration of note is that made by Dechand \textit{et al.} \cite{dechand2016empirical} where they briefly consider medical conditions such as ADHD and reading or visual disorders and the way they affect the comparison's effectiveness. They highlight a slight reduction in overall accuracy, although due to their small sample size, they cannot conclusively validate these results. This is an aspect unique to all literature. Further work would be required to produce conclusive results. Therefore, highlighting a gap in the literature.

One glaring issue with the demographical health of \textbf{Ersin Uzun \textit{et al.}}\cite{uzun2007usability} study is the use of two entirely different groups of participants. Not only were their demographics different, but they were from different countries (America and Finland). Different cultures contain inherent biases and assumptions. Moreover, more concerns are raised around the dual study design of this paper with two completely different set of participants. Changes were made pragmatically regarding the results from the first round of 40 participants. These were then re-assessed and the results were directly compared. This puts huge doubts on the validity of the results with no consideration made by the authors to control external factors that may affect the performance of the method of comparison.

Hexadecimal and numerical schemes were not included as encoding schemes in work by \textbf{Hsiao, \textit{et al.}}\cite{hsiao2009study} (some of the most widespread encoding schemes). This decision was based on their claims on similarities to Base32 and ``well-known deficiencies" in the excluded schemes. This point was provided with no further justification or quantified in any way. It is also not consistent with available research, for example, in \textit{``Empirical Study of Textual Key-Fingerprint Representations''}\cite{dechand2016empirical} it was shown numerical representations performed significantly better than that of Base32.

\begin{table}[h!]
    \makebox[\textwidth][c]{
        \input{tables/fingerprint_attacker_stren.tex}
    }%
    \caption{Paper attribute comparison}
    \label{tab:attacker}
\end{table}

Table \ref{tab:attacker} has been provided to visually compare the differing aspects of the papers' parameters. Clearly it can be seen from the table the large ranges in participant size, entropy and attacker strength.

%TODO: Maybe talk about the effect of visual or mental impairments 
\subsubsection{Topic conclusion}
Overall, this review has identified several key areas suitable for further work.
The first is the performance assessment of graphical encoding schemes. Recreation of pre-existing results from \cite{hsiao2009study}\cite{tan2017can} is required to corroborate current conclusions and validate results.
Another gap in the research is the consideration into the utilization of encoding schemes in realistic conditions, i.e. ``remove vs. proximity." This topic was initially covered by \cite{shirvanian2017pitfalls} but their scope was limited. Further work, therefore, could increased the scope and touch upon a large number of schemes in these settings.
The final aspect for further work is the limited consideration into the feasibility of attacks on encoding schemes. All of the papers assessed simulated attacks and had minimal consideration for the execution of these attacks. Therefore, further work could delve into the implementation of such attacks and their feasibility in terms of compute and storage complexity.

\subsection{Encoding schemes}
Another area of research is investigations into the actual physical encodings of the hash digest. This section will briefly discuss the current research available on the creation and security of actual encoding schemes. The actual details of the operation of the schemes are outside the scope of this literature review, therefore, minimal attention will be allocated to these details.

@@

\subsection{Attacks on encoding schemes}
This area of research studies ways to physically execute attacks on fingerprint encoding schemes. This differs from previously examined work due to papers discussing the performance and fallibility of encoding schemes simulated the attack without consideration for how the attack would be performed. Research in this area is scant, with lots of research attention being directed towards the security of Man-in-the-Middle (MITM) attacks and not the encoding schemes themselves.
